### 2021-03-05 paper3/paper3.tex
Use British spelling

%\icmltitle{Learning with Perfect Bags:\\
%Addressing Hidden Stratification with Zero Labelled Data}
\textsc{Authors}:\\
Thomas Kehrenberg$^1$, Viktoriia Sharmanska$^1$, Myles Bartlett$^1$, and Novi Quadrianto$^1$ \\
\textsc{Affiliations}:\\
$^1$ Predictive Analytics Lab (PAL), University of Sussex, Brighton, UK\\
\textsc{Conference}:\;\; Submitted to \textit{Internation Conference on Machine Learning} (ICML), 2021 \\
% \textsc{DOI}:\;\;10.3389/frai.2020.00033
\textsc{Note}:\;\; The appendix has been included as section~\ref{sec:zsf-appendix}.

@@

Algorithms such as classifiers, however, are trained from large amount of labelled data, and are typically trained to optimise \emph{global} metrics such as average classification accuracy.
%
% \begin{figure*}[t]
% \centering
% \includegraphics[width=0.6\textwidth,page=2]{figures/ideal.pdf}
% \caption{Overview of the zero-shot stratification problem. Training dataset (pairs of input data $x$ and class label $y$) will only contain data points that are labeled ``yes'' by the decision policy. This systematic bias might result in a subgroup(s) to have zero labeled data (in the example above, ``married'' subgroup has zero labeled data). The subgrouping information $s$ is unavailable/unlabeled at deployment time, and is only partially labeled at train time.}
% \vspace{-0.5cm}
% \label{fig:censoring}
% \end{figure*}
In many real-world classification tasks, each labelled class consists of multiple semantically distinct subclasses, or subgroups.
For example, the ``dog'' class label can have finer-grained intra-class variations, such as ``dog indoor'' and ``dog outdoor''.
%
This finer-grained subgrouping information is typically unavailable/unlabeled (e.g. \citep{SohDunAngGuetal20,nam2020learning}).
%
The standard training process brings about two inter-connected challenges: a) classifiers often under-perform on important hidden subgroups (\emph{hidden stratification}) \citep{RayDunCarRe20,SohDunAngGuetal20};
and b) \emph{systematic bias}~\citep{kallus2018residual} affects whether or not entire collections of data points appear in the training dataset,
and can make the classifier unprepared for treating those subgroups in the eventual deployment setting
% and can hamper attempts to correct for bias in the eventual deployment data
(\emph{residual bias}). %\cite{kallus2018residual}.

We are interested in hidden stratification in which, for some of the subgroups, labelled training data is only available with a certain outcome, or labelled training samples are not available at all, due to systematic bias.
This can be seen as a strong sampling bias.
%
For instance, data on loan defaults can only be collected on those loan
applicants who were approved in the past \citep{kallus2018residual}.
%
Here a loan decision policy specifies whether an individual will be included in the training dataset.
%
Individuals can be thought of as belonging to a specific subgroup such as ``married'' or ``not married'' and systematic bias produced by a historical decision policy may result in the ``not married'' subgroup having poor, or altogether non-existent, representation.
%For instance, a medical imaging model trained to classify between ``benign'' and ``abnormal'' lesions may achieve high overall performance, yet consistently mislabel a rare but critical abnormal subgroup as ``benign'' \cite{DunYiLanReetal19}.
%
%Each class actually composes of ``no-patch'' and ``patch'' hidden subgroups, and the ``no-patch'' subgroup is further splitted into “histopathology” and “non-histopathology”.
%
%A past decision policy of \emph{not} requiring histopathology (biopsy \& pathologist referral) specifies whether an individual will be included in the training dataset.
%
%This systematic bias might lead to the no-patch subgroup to have thin or non-existent labeled data.
%
We formalise this stratification problem as a data setting where a decision policy can lead to one or more subgroups having no labelled data.

@@

\paragraph{Disentanglement guarantees.}
%\citet{ShuCheKumErmetal20} decomposed disentanglement into two distinct concepts: consistency and restrictiveness.
%
%The decomposition allows for correlated factors of variations, which is crucial in our case as the subgroup label $s$ is correlated with the class label $y$. 
%
%We now provide a single-factor disentanglement guarantee of our disentanglement with a collection of perfect bags approach based on theoretical results of \citet{ShuCheKumErmetal20}.
%
Under \citet{ShuCheKumErmetal20}'s framework, our disentanglement supervision corresponds to \emph{match pairing}:
We observe samples from a pair of distributions (deployment set and training set).
As an example, we consider colored MNIST with two classes (digits 2 and 4) and two subgroups (colors purple and green), where the training set is lacking the combination of digit 4 and color purple.
Sample pairs from the two distributions have a shared underlying factor: 2's can be purple and green.
The pairs also have a factor that differs: in the training set, a 4 can only be green; in the deployment set, it can have both colors.
Theorem 1 in \citep{ShuCheKumErmetal20} guarantees then that a (sufficiently powerful) disentangling encoder will produce a disentangling that is \emph{consistent} with respect to the shared factor and \emph{restricted} with respect to the changing factor.
In our case, $z_y$ will capture the shared factor and $z_s$ the changing factor.
For the invariant representation, we drop $z_s$, which can be seen as setting this value to 0; as $z_s$ has guaranteed \emph{restrictiveness}, the effect of changing $z_s$ is restricted to changing the corresponding factor,
which in the example refers to whether or not 4's can be purple.
%We show that this is in fact the case for our approach, and we focus on the $z_s$ factor for the explanation.
%
%Share pairing is where $z_s$ is the only factor that is shared between training bags and perfect bags\footnote{We focus on groups of data points in contrast to the paired setting in \citet{ShuCheKumErmetal20}.}. 
%
%Change pairing is where $z_s$ is the only factor that is changed.
%

@@

    Results from 30 repeats for the Colored MNIST dataset with two digits, 2 and 4, with \emph{subgroup bias} for the color `{\color{purple}purple}': for {\color{purple}purple}, only the digit class `2' is present.

@@

    Results from 30 repeats for the Colored MNIST dataset with two digits, 2 and 4, with a \emph{missing subgroup}: the training dataset only has {\color{green}green} digits.

@@

    Results from 30 repeats for the Colored MNIST dataset with three digits: `2', `4' and `6'.
    Four combinations of digit and color are missing: {\color{green}green} 2's, {\color{blue}blue} 2's, {\color{blue}blue} 4's and {\color{green}green} 6's.

@@

\subsection{Implementation}
\label{sec:implementation}
Our proposed framework, demonstrated in fig.\,\ref{fig:architecture}, entails two steps: 1) sample perfect bags from an unlabeled deployment set, and 2) produce disentangled representations using perfect bags for adversarial distribution-matching.

\paragraph{Constructing approximately perfect bags via clustering.}
We cluster the data points from the deployment set into $K=\textrm{dim}(\mathcal{Y})\cdot \textrm{dim}(\mathcal{S})$ number of clusters, i.e. the number of data sources $\Omega_{y,s}$. 
%, using a learned embedding $z$.
%
We use the k-means clustering algorithm, and a recently proposed method based on rank statistics \citep{HanRebEhrVedetal20}.
%
The cluster assignments can then be used to evenly stratify the deployment set into perfect batches, to be used by the subsequent disentangling phase.
%
% The objective for clustering can be written as a binary cross-entropy loss, induced by the following binary classification problem -- for any two data points in the deployment set to decide whether or not they belong to the same cluster:
% %
% \begin{align}
%  \mathcal{L}_{\textit{unsup}} &= -\frac{1}{N^2}\sum_{i,j} [s_{ij}\: \text{log} (C(z_i)^\top C(z_j)) \nonumber\\
%  &+ (1-s_{ij}) \: \text{log}(1-C(z_i)^\top C(z_j))].
%  \label{eq:clustering}
%  \end{align}
% %
% Here, $N^2$ is the amount of sample pairs,  $s_{ij}$ are binary labels capturing that similar data points belong to the same group, $s_{ij}=1$, while dissimilar ones do not, $s_{ij}=0$. 
% %
% Since the deployment set is unlabeled, $s_{ij}$ are unknown. 
% %
% We estimate $s_{ij}$ with the robust rank statistics approach (described below) and use the estimates as pseudo-labels in \eqref{eq:clustering}. 
% %
% The dot-product $C(z_i)^\top C(z_j)$ measures the similarity of the two data points $z_i$ and $z_j$ to score whether they share a cluster.
% %
% This is computed on the embedding of the softmax classifier $C: \mathcal{Z} \rightarrow \{1,\ldots, K\}$. 
% %
% In addition to this unsupervised training, the classifier $C$ is also trained on the labeled data using the standard cross-entropy loss.  
% %We train the classifier $C$ on the training data using a standard cross-entropy loss and the labeled sources $\Omega_{y,s}$. 
% %
% This way, the data points in the deployment set that are similar to the known demographics are assigned to the associated clusters, while the dissimilar data points form the new clusters corresponding to the invisible groups.

% The method of rank statistics \citep{YagStrRosLin11,DeaRuzSegShletal13,HanRebEhrVedetal20} avoids comparing the vectors $z_i$ and $z_j$ directly, and instead ranks the values in $z_i$ and in $z_j$ by their magnitudes, and compares the top-$U$ rankings (we found $U = 5$ to perform well in our experiments).
% %
% If the rankings obtained for two unlabeled data points $z_i$ and $z_j$ are the same, they are assigned to the same class, $s_{ij}=1$. 
% %
As a result of clustering, the data points in the deployment set $\gD_\mathit{dep}$ are labeled with cluster assignments 
$\gD_\mathit{dep}=\{(x_i, c_i)\}$, $c_i = C(z_i)$.
%
We balance $\gD_\mathit{dep}$ so that all clusters have equal size to form a perfect bag
% (see fig. \ref{fig:teaser})
, and use it as a supervision signal for the disentangling step.
%
%Note that our method of constructing the perfect dataset based on clustering does not require solving the linear assignment problem of cluster-source association. This is particularly suitable for exploring the setting where invisible demographics are formed of multiple sources.

@@

\paragraph{Clustering guarantees.}
By leveraging the feature space of deep models, previous research has shown that semantically meaningful clusters can be discovered without the need for ground-truth annotations
 (for example \citet{RayDunCarRe20,GanVanGeoProetal20,SohDunAngGuetal20,HanRebEhrVedetal20}). 
\citet{SohDunAngGuetal20} recently provided a ``clustering guarantee'' when assuming access to ground truth class labels $y$ and using clustering within each class to generate approximate subgroups $s$.  
%
Even with infinite data, accurate identification of the subgroups is unfortunately impossible. 
%
Instead, \citet{SohDunAngGuetal20} showed that the error in their quantity of interest, classification loss for each subgroup, can be bounded by the total variation estimation error in the mixture-of-Gaussians case.   
%
Our primary goal is to ensure good construction of perfect bags.
We can adapt the analysis of \citet{SohDunAngGuetal20} to bound the error of our quantity of interest, \emph{aggregate statistics} of the approximated perfect bag for each subgroup.
%
We can subsequently apply the union bound to take into account that we have a fully-unlabeled setting.

\subsection{Limitation and intended use}
\label{sec:limitations}
% First, dataset consumers should take extra care about the cost-benefit analysis of selecting particular datasets for their machine learning tasks. 
%
Although having zero labeled examples for some subgroups is not uncommon due to the effects of systematic bias,
we should make a value-judgment on the efficacy of the dataset with respect to a task.
%
% {\color{red}Corrective action such as the one described in this paper or inaction should be recorded.}
We can then decide whether or not to take corrective action as described in this paper.

@@

% In this paper, we aim to highlight the problem of zero-label stratification that can result from systematic bias, and to make a case for it as something to be seriously considered (even if to be dismissed) when planning, building, evaluating and regulating machine learning systems.
%Clearly, the use of algorithmic approaches still requires human review in a manner that is similar to error auditing, but is less dependent on the specific human auditor to initially identify the stratifica- tion. While algorithmic approaches to measurement can reduce burden on human analysts and take advantage of learned encodings to identify subclasses, their efficacy is limited by the difficulty of separating important subclasses in the feature space analyzed.

\section{Experiments}
We perform experiments using image and tabular datasets: Colored MNIST \citep{kim2019learning}, CelebA \citep{liu2015faceattributes} and Adult Income\footnote{{Results for Adult Income dataset and discussion can be found in the supplementary material.}} \citep{Dua:2017} that are publicly available. 
%arjovsky2019invariant, KehBarThoQua20
%
To validate the first step of creating the perfect bags, we compare the performance of our disentangling model when paired with each of four different balancing methods:
1) with clustering via rank statistics (\texttt{Ranking});
2) with clustering via k-means (\texttt{K-means}); 
3) without balancing, when the deployment set $\gD_{dep}$ is used as is (\texttt{No bal.});
4) with balancing done using the ground-truth class and subgroup labels (\texttt{Perfect}) that would in practice be unobservable; this provides insight into what can be achieved under ideal conditions and how sensitive the method is to imperfections in the bag-balancing.

@@

%
%%%%%OR
%During training, the {\color{green}green} color group is observed in both digits (sources $\Omega_{y=\text{'two' }s=\color{green}green}$ and $\Omega_{y=\text{'four' }s=\color{green}green}$), but for the {\color{purple}purple} color group, we only observe the one-sided outcome ($\Omega_{y=\text{'two'}s=\color{purple}purple}$), so the source $\Omega_{y=\text{'four'}s=\color{green}green}$ defines invisible demographic.  
%
\subsection{Colored MNIST}
The MNIST dataset \citep{lecun1998gradient} consists of 70,000 (60,000 designated for training, 10,000 for testing) images of gray-scale hand-written digits. We color the digits following a similar procedure to that outlined by \citep{kim2019learning}, randomly assigning each sample one of ten distinct RGB colors. Each source is then a combination of digit-class (class label) and color (subgroup label). We use no data-augmentation aside from symmetrically zero-padding the images to be of size 32x32.
We create imbalance in both $D_{dep}$ and $D_{tr}$ by sub-sampling the remaining sources;
we do this to render a more realistic setting in which the deployment set is not innately balanced and demands preliminary clustering to construct approximately perfect bags. The sub-sampling proportions used for each set of experiments can be found in the appendix.

We begin by considering a binary, 2-digit, 2-color, variant of the dataset with $Y = \{2, 4\}$ and $S = \{\text{\color{green}green}, \text{\color{purple}purple}\}$.
For this variant we explore both the SB (subgroup bias) and MS (missing subgroup) settings.
To simulate the SB setting, we set  $\Omega_{y=4,s=\text{\color{purple}{purple}}}$ to be the missing source.
To simulate the MS setting, where we have training data for only a single subgroup, we set both $\Omega_{y=\text{'two'},s=\color{purple}purple}$ and $\Omega_{y=\text{'four'},s=\color{purple}purple}$ to be the missing sources (i.e. $D_{tr}$ consists of only $\text{\color{green}{green}}$ digits).
%

@@

To investigate how an increase in the number of classes affects disentangling of classes and groups, we look to a 3-digit, 3-color variant of the dataset in the SB setting where four sources are missing from $D_{tr}$.
Results for this configuration are shown in fig.~\ref{fig:cmnist-3dig-4miss}.
We see that the performance of \texttt{No bal.} is quite close to that of \texttt{Perfect}, we suspect this is because balancing is less critical with the increased number of subgroups strengthening the training signal.
As the PR ratio is not a suitable metric for non-binary $S$,
we instead quantify the invariance of the predictions to the subgroup with
the HGR maximal correlation \citep{renyi1959measures}.

\subsection{CelebA}
To demonstrate our method can be used to mitigate biased decision making for real-world computer vision problems, we consider the CelebA dataset~\citep{liu2015faceattributes} comprising over 200,000 images of different celebrities.
The dataset comes with per-image annotations of attributes related to visual appearance, emotion, gender, age.
%
We predict the smiling attribute as the class label and use the binary attribute, gender, as the subgroup label.
Here, we consider an SB setting, where ``smiling females'', $\Omega_{s=0, y=1}$, constitutes a missing source.
%
Since the dataset exhibits natural imbalance in $S \times Y$, we perform no additional sub-sampling of either the training set or the deployment set, as we did for Colored MNIST.
%
Fig.~\ref{fig:celeba-gender-smiling} shows the model trained with a perfectly balanced deployment set \texttt{Perfect} outperforms \texttt{No bal.}, indicating that this natural imbalance is sufficiently strong to somewhat disrupt the disentangling procedure and is something to be potentially remedied through clustering.
%
We did not use the clustering approach for CelebA, as the attributes ``smiling'' and ``gender'' are not the most salient;
more work is needed on the clustering side to discover all semantically meaningful clusters.
%
Nonetheless, our method, with or without the artificial balancing, yields much better accuracy than that of \texttt{ERM}. Furthermore, we show qualitative results of the disentangling in fig.~\ref{fig:celeba-recons}, and note a clear separation of subgroup-relevant information from subgroup-irrelevant information.
%in $g(0, f_s(x))$ and $g(f_y(x), 0)$, respectively.

>>>>

%\icmltitle{Learning with Perfect Bags:\\
%Addressing Hidden Stratification with Zero Labelled Data}
\textsc{Authors}:\\
Thomas Kehrenberg$^1$, Viktoriia Sharmanska$^1$, Myles Bartlett$^1$, and Novi Quadrianto$^1$ \\
\textsc{Affiliations}:\\
$^1$ Predictive Analytics Lab (PAL), University of Sussex, Brighton, UK\\
\textsc{Conference}:\;\; Submitted to \textit{International Conference on Machine Learning} (ICML), 2021 \\
% \textsc{DOI}:\;\;10.3389/frai.2020.00033
\textsc{Note}:\;\; The appendix has been included as section~\ref{sec:zsf-appendix}.

@@

Algorithms such as classifiers, however, are trained from large amount of labelled data, and are typically trained to optimise \emph{global} metrics such as average classification accuracy.
%
% \begin{figure*}[t]
% \centering
% \includegraphics[width=0.6\textwidth,page=2]{figures/ideal.pdf}
% \caption{Overview of the zero-shot stratification problem. Training dataset (pairs of input data $x$ and class label $y$) will only contain data points that are labelled ``yes'' by the decision policy. This systematic bias might result in a subgroup(s) to have zero labelled data (in the example above, ``married'' subgroup has zero labelled data). The subgrouping information $s$ is unavailable/unlabelled at deployment time, and is only partially labelled at train time.}
% \vspace{-0.5cm}
% \label{fig:censoring}
% \end{figure*}
In many real-world classification tasks, each labelled class consists of multiple semantically distinct subclasses, or subgroups.
For example, the ``dog'' class label can have finer-grained intra-class variations, such as ``dog indoor'' and ``dog outdoor''.
%
This finer-grained subgrouping information is typically unavailable/unlabelled (e.g. \citep{SohDunAngGuetal20,nam2020learning}).
%
The standard training process brings about two inter-connected challenges: a) classifiers often under-perform on important hidden subgroups (\emph{hidden stratification}) \citep{RayDunCarRe20,SohDunAngGuetal20};
and b) \emph{systematic bias}~\citep{kallus2018residual} affects whether or not entire collections of data points appear in the training dataset,
and can make the classifier unprepared for treating those subgroups in the eventual deployment setting
% and can hamper attempts to correct for bias in the eventual deployment data
(\emph{residual bias}). %\cite{kallus2018residual}.

We are interested in hidden stratification in which, for some of the subgroups, labelled training data is only available with a certain outcome, or labelled training samples are not available at all, due to systematic bias.
This can be seen as a strong sampling bias.
%
For instance, data on loan defaults can only be collected on those loan
applicants who were approved in the past \citep{kallus2018residual}.
%
Here a loan decision policy specifies whether an individual will be included in the training dataset.
%
Individuals can be thought of as belonging to a specific subgroup such as ``married'' or ``not married'' and systematic bias produced by a historical decision policy may result in the ``not married'' subgroup having poor, or altogether non-existent, representation.
%For instance, a medical imaging model trained to classify between ``benign'' and ``abnormal'' lesions may achieve high overall performance, yet consistently mislabel a rare but critical abnormal subgroup as ``benign'' \cite{DunYiLanReetal19}.
%
%Each class actually composes of ``no-patch'' and ``patch'' hidden subgroups, and the ``no-patch'' subgroup is further splitted into “histopathology” and “non-histopathology”.
%
%A past decision policy of \emph{not} requiring histopathology (biopsy \& pathologist referral) specifies whether an individual will be included in the training dataset.
%
%This systematic bias might lead to the no-patch subgroup to have thin or non-existent labelled data.
%
We formalise this stratification problem as a data setting where a decision policy can lead to one or more subgroups having no labelled data.

@@

\paragraph{Disentanglement guarantees.}
%\citet{ShuCheKumErmetal20} decomposed disentanglement into two distinct concepts: consistency and restrictiveness.
%
%The decomposition allows for correlated factors of variations, which is crucial in our case as the subgroup label $s$ is correlated with the class label $y$. 
%
%We now provide a single-factor disentanglement guarantee of our disentanglement with a collection of perfect bags approach based on theoretical results of \citet{ShuCheKumErmetal20}.
%
Under \citet{ShuCheKumErmetal20}'s framework, our disentanglement supervision corresponds to \emph{match pairing}:
We observe samples from a pair of distributions (deployment set and training set).
As an example, we consider coloured MNIST with two classes (digits 2 and 4) and two subgroups (colours purple and green), where the training set is lacking the combination of digit 4 and colour purple.
Sample pairs from the two distributions have a shared underlying factor: 2's can be purple and green.
The pairs also have a factor that differs: in the training set, a 4 can only be green; in the deployment set, it can have both colours.
Theorem 1 in \citep{ShuCheKumErmetal20} guarantees then that a (sufficiently powerful) disentangling encoder will produce a disentangling that is \emph{consistent} with respect to the shared factor and \emph{restricted} with respect to the changing factor.
In our case, $z_y$ will capture the shared factor and $z_s$ the changing factor.
For the invariant representation, we drop $z_s$, which can be seen as setting this value to 0; as $z_s$ has guaranteed \emph{restrictiveness}, the effect of changing $z_s$ is restricted to changing the corresponding factor,
which in the example refers to whether or not 4's can be purple.
%We show that this is in fact the case for our approach, and we focus on the $z_s$ factor for the explanation.
%
%Share pairing is where $z_s$ is the only factor that is shared between training bags and perfect bags\footnote{We focus on groups of data points in contrast to the paired setting in \citet{ShuCheKumErmetal20}.}. 
%
%Change pairing is where $z_s$ is the only factor that is changed.
%

@@

    Results from 30 repeats for the Coloured MNIST dataset with two digits, 2 and 4, with \emph{subgroup bias} for the colour `{\color{purple}purple}': for {\color{purple}purple}, only the digit class `2' is present.

@@

    Results from 30 repeats for the Coloured MNIST dataset with two digits, 2 and 4, with a \emph{missing subgroup}: the training dataset only has {\color{green}green} digits.

@@

    Results from 30 repeats for the Coloured MNIST dataset with three digits: `2', `4' and `6'.
    Four combinations of digit and colour are missing: {\color{green}green} 2's, {\color{blue}blue} 2's, {\color{blue}blue} 4's and {\color{green}green} 6's.

@@

\subsection{Implementation}
\label{sec:implementation}
Our proposed framework, demonstrated in fig.\,\ref{fig:architecture}, entails two steps: 1) sample perfect bags from an unlabelled deployment set, and 2) produce disentangled representations using perfect bags for adversarial distribution-matching.

\paragraph{Constructing approximately perfect bags via clustering.}
We cluster the data points from the deployment set into $K=\textrm{dim}(\mathcal{Y})\cdot \textrm{dim}(\mathcal{S})$ number of clusters, i.e. the number of data sources $\Omega_{y,s}$. 
%, using a learned embedding $z$.
%
We use the k-means clustering algorithm, and a recently proposed method based on rank statistics \citep{HanRebEhrVedetal20}.
%
The cluster assignments can then be used to evenly stratify the deployment set into perfect batches, to be used by the subsequent disentangling phase.
%
% The objective for clustering can be written as a binary cross-entropy loss, induced by the following binary classification problem -- for any two data points in the deployment set to decide whether or not they belong to the same cluster:
% %
% \begin{align}
%  \mathcal{L}_{\textit{unsup}} &= -\frac{1}{N^2}\sum_{i,j} [s_{ij}\: \text{log} (C(z_i)^\top C(z_j)) \nonumber\\
%  &+ (1-s_{ij}) \: \text{log}(1-C(z_i)^\top C(z_j))].
%  \label{eq:clustering}
%  \end{align}
% %
% Here, $N^2$ is the amount of sample pairs,  $s_{ij}$ are binary labels capturing that similar data points belong to the same group, $s_{ij}=1$, while dissimilar ones do not, $s_{ij}=0$. 
% %
% Since the deployment set is unlabelled, $s_{ij}$ are unknown. 
% %
% We estimate $s_{ij}$ with the robust rank statistics approach (described below) and use the estimates as pseudo-labels in \eqref{eq:clustering}. 
% %
% The dot-product $C(z_i)^\top C(z_j)$ measures the similarity of the two data points $z_i$ and $z_j$ to score whether they share a cluster.
% %
% This is computed on the embedding of the softmax classifier $C: \mathcal{Z} \rightarrow \{1,\ldots, K\}$. 
% %
% In addition to this unsupervised training, the classifier $C$ is also trained on the labeled data using the standard cross-entropy loss.  
% %We train the classifier $C$ on the training data using a standard cross-entropy loss and the labeled sources $\Omega_{y,s}$. 
% %
% This way, the data points in the deployment set that are similar to the known demographics are assigned to the associated clusters, while the dissimilar data points form the new clusters corresponding to the invisible groups.

% The method of rank statistics \citep{YagStrRosLin11,DeaRuzSegShletal13,HanRebEhrVedetal20} avoids comparing the vectors $z_i$ and $z_j$ directly, and instead ranks the values in $z_i$ and in $z_j$ by their magnitudes, and compares the top-$U$ rankings (we found $U = 5$ to perform well in our experiments).
% %
% If the rankings obtained for two unlabeled data points $z_i$ and $z_j$ are the same, they are assigned to the same class, $s_{ij}=1$. 
% %
As a result of clustering, the data points in the deployment set $\gD_\mathit{dep}$ are labelled with cluster assignments 
$\gD_\mathit{dep}=\{(x_i, c_i)\}$, $c_i = C(z_i)$.
%
We balance $\gD_\mathit{dep}$ so that all clusters have equal size to form a perfect bag
% (see fig. \ref{fig:teaser})
, and use it as a supervision signal for the disentangling step.
%
%Note that our method of constructing the perfect dataset based on clustering does not require solving the linear assignment problem of cluster-source association. This is particularly suitable for exploring the setting where invisible demographics are formed of multiple sources.

@@

\paragraph{Clustering guarantees.}
By leveraging the feature space of deep models, previous research has shown that semantically meaningful clusters can be discovered without the need for ground-truth annotations
 (for example \citet{RayDunCarRe20,GanVanGeoProetal20,SohDunAngGuetal20,HanRebEhrVedetal20}). 
\citet{SohDunAngGuetal20} recently provided a ``clustering guarantee'' when assuming access to ground truth class labels $y$ and using clustering within each class to generate approximate subgroups $s$.  
%
Even with infinite data, accurate identification of the subgroups is unfortunately impossible. 
%
Instead, \citet{SohDunAngGuetal20} showed that the error in their quantity of interest, classification loss for each subgroup, can be bounded by the total variation estimation error in the mixture-of-Gaussians case.   
%
Our primary goal is to ensure good construction of perfect bags.
We can adapt the analysis of \citet{SohDunAngGuetal20} to bound the error of our quantity of interest, \emph{aggregate statistics} of the approximated perfect bag for each subgroup.
%
We can subsequently apply the union bound to take into account that we have a fully-unlabelled setting.

\subsection{Limitation and intended use}
\label{sec:limitations}
% First, dataset consumers should take extra care about the cost-benefit analysis of selecting particular datasets for their machine learning tasks. 
%
Although having zero labelled examples for some subgroups is not uncommon due to the effects of systematic bias,
we should make a value-judgement on the efficacy of the dataset with respect to a task.
%
% {\color{red}Corrective action such as the one described in this paper or inaction should be recorded.}
We can then decide whether or not to take corrective action as described in this paper.

@@

% In this paper, we aim to highlight the problem of zero-label stratification that can result from systematic bias, and to make a case for it as something to be seriously considered (even if to be dismissed) when planning, building, evaluating and regulating machine learning systems.
%Clearly, the use of algorithmic approaches still requires human review in a manner that is similar to error auditing, but is less dependent on the specific human auditor to initially identify the stratifica- tion. While algorithmic approaches to measurement can reduce burden on human analysts and take advantage of learned encodings to identify subclasses, their efficacy is limited by the difficulty of separating important subclasses in the feature space analysed.

\section{Experiments}
We perform experiments using image and tabular datasets: Coloured MNIST \citep{kim2019learning}, CelebA \citep{liu2015faceattributes} and Adult Income\footnote{{Results for Adult Income dataset and discussion can be found in the supplementary material.}} \citep{Dua:2017} that are publicly available. 
%arjovsky2019invariant, KehBarThoQua20
%
To validate the first step of creating the perfect bags, we compare the performance of our disentangling model when paired with each of four different balancing methods:
1) with clustering via rank statistics (\texttt{Ranking});
2) with clustering via k-means (\texttt{K-means}); 
3) without balancing, when the deployment set $\gD_{dep}$ is used as is (\texttt{No bal.});
4) with balancing done using the ground-truth class and subgroup labels (\texttt{Perfect}) that would in practice be unobservable; this provides insight into what can be achieved under ideal conditions and how sensitive the method is to imperfections in the bag-balancing.

@@

%
%%%%%OR
%During training, the {\color{green}green} color group is observed in both digits (sources $\Omega_{y=\text{'two' }s=\color{green}green}$ and $\Omega_{y=\text{'four' }s=\color{green}green}$), but for the {\color{purple}purple} color group, we only observe the one-sided outcome ($\Omega_{y=\text{'two'}s=\color{purple}purple}$), so the source $\Omega_{y=\text{'four'}s=\color{green}green}$ defines invisible demographic.  
%
\subsection{Coloured MNIST}
The MNIST dataset \citep{lecun1998gradient} consists of 70,000 (60,000 designated for training, 10,000 for testing) images of grey-scale hand-written digits. We colour the digits following a similar procedure to that outlined by \citep{kim2019learning}, randomly assigning each sample one of ten distinct RGB colours. Each source is then a combination of digit-class (class label) and colour (subgroup label). We use no data-augmentation aside from symmetrically zero-padding the images to be of size 32x32.
We create imbalance in both $D_{dep}$ and $D_{tr}$ by sub-sampling the remaining sources;
we do this to render a more realistic setting in which the deployment set is not innately balanced and demands preliminary clustering to construct approximately perfect bags. The sub-sampling proportions used for each set of experiments can be found in the appendix.

We begin by considering a binary, 2-digit, 2-colour, variant of the dataset with $Y = \{2, 4\}$ and $S = \{\text{\color{green}green}, \text{\color{purple}purple}\}$.
For this variant we explore both the SB (subgroup bias) and MS (missing subgroup) settings.
To simulate the SB setting, we set  $\Omega_{y=4,s=\text{\color{purple}{purple}}}$ to be the missing source.
To simulate the MS setting, where we have training data for only a single subgroup, we set both $\Omega_{y=\text{'two'},s=\color{purple}purple}$ and $\Omega_{y=\text{'four'},s=\color{purple}purple}$ to be the missing sources (i.e. $D_{tr}$ consists of only $\text{\color{green}{green}}$ digits).
%

@@

To investigate how an increase in the number of classes affects disentangling of classes and groups, we look to a 3-digit, 3-colour variant of the dataset in the SB setting where four sources are missing from $D_{tr}$.
Results for this configuration are shown in fig.~\ref{fig:cmnist-3dig-4miss}.
We see that the performance of \texttt{No bal.} is quite close to that of \texttt{Perfect}, we suspect this is because balancing is less critical with the increased number of subgroups strengthening the training signal.
As the PR ratio is not a suitable metric for non-binary $S$,
we instead quantify the invariance of the predictions to the subgroup with
the HGR maximal correlation \citep{renyi1959measures}.

\subsection{CelebA}
To demonstrate our method can be used to mitigate biased decision making for real-world computer vision problems, we consider the CelebA dataset~\citep{liu2015faceattributes} comprising over 200,000 images of different celebrities.
The dataset comes with per-image annotations of attributes related to visual appearance, emotion, gender, age.
%
We predict the smiling attribute as the class label and use the binary attribute, gender, as the subgroup label.
Here, we consider an SB setting, where ``smiling females'', $\Omega_{s=0, y=1}$, constitutes a missing source.
%
Since the dataset exhibits natural imbalance in $S \times Y$, we perform no additional sub-sampling of either the training set or the deployment set, as we did for Coloured MNIST.
%
Fig.~\ref{fig:celeba-gender-smiling} shows the model trained with a perfectly balanced deployment set \texttt{Perfect} outperforms \texttt{No bal.}, indicating that this natural imbalance is sufficiently strong to somewhat disrupt the disentangling procedure and is something to be potentially remedied through clustering.
%
We did not use the clustering approach for CelebA, as the attributes ``smiling'' and ``gender'' are not the most salient;
more work is needed on the clustering side to discover all semantically meaningful clusters.
%
Nonetheless, our method, with or without the artificial balancing, yields much better accuracy than that of \texttt{ERM}. Furthermore, we show qualitative results of the disentangling in fig.~\ref{fig:celeba-recons}, and note a clear separation of subgroup-relevant information from subgroup-irrelevant information.
%in $g(0, f_s(x))$ and $g(f_y(x), 0)$, respectively.
